Stock enhancement undermines community stability

### 2022-02-04 rmd/manuscript.Rmd
edit: manuscript

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the demographic benefit of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across most ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, $\sigma_{\epsilon}$, and $f_R$; **Figure S1-6**).

@@

Stock enhancement, however, had little influence on community dynamics when carrying capacity was sufficiently large ($K=500$; **Figures S7-S12**). In particular, stock enhancement increased the enhanced species with a low population growth rate ($r_1=0.5$; **Figures S7 and S8**). This result may explain why some of the best evidence for successful stock enhancement comes from long-lived endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. More importantly, the contrasting community response at different carrying capacities provides deeper insights. One of the core motivations for stock enhancement is to mitigate declining trends of natural resources due to human impacts, such as habitat loss [@chaseEcosystemDecayExacerbates2020]. Ironically, our results suggest that stock enhancement will not bring desired outcomes unless we resolve root causes that compromise environmental capacity first.

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the stock enhancement of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (**Figure 1C and 1D**), a long-term program exists to monitor stream fish communities along with the official stocking records. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see **Materials and Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with intensive stocking (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest stocking level (**Figure 1B**). Our analysis strongly supported the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S8**). This pattern was associated with the reduced long-term average of the total community density and fewer species richness (**Figure 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figure 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with stock enhancement (**Figure 1B and 2**).

>>>>

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the reproductive contribution of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across most ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, $\sigma_{\epsilon}$, and $f_R$; **Figures S1-S3**).

@@

Stock enhancement, however, had little influence on community dynamics when carrying capacity was sufficiently large ($K=500$; **Figures S4-S6**). In particular, stock enhancement increased the enhanced species with a low population growth rate ($r_1=0.5$; **Figures S4 and S5**). This result may explain why some of the best evidence for successful stock enhancement comes from long-lived endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. More importantly, the contrasting community response at different carrying capacities provides deeper insights. One of the core motivations for stock enhancement is to mitigate declining trends of natural resources due to human impacts, such as habitat loss [@chaseEcosystemDecayExacerbates2020]. Ironically, our results suggest that stock enhancement will not bring desired outcomes unless we resolve root causes that compromise environmental capacity first.

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the stock enhancement of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (**Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official stocking records. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see **Materials and Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with intensive stocking (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest stocking level (**Figure 1B**). Our analysis strongly supported the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S8**). This pattern was associated with the reduced long-term average of the total community density and fewer species richness (**Figure 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with stock enhancement (**Figures 1B and 2**).
